An open-source integrated framework for the automation of citation collection and screening in systematic reviews.

### 2022-03-03 Manuscript/Manuscript.tex
prepare BMC submission

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}        % https://github.com/rstudio/rticles/issues/343
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{graphicx}

@@

\author{
    Angelo D'Ambrosio
   \\
    Institute for Infection Prevention and Hospital Hygiene \\
    Freiburg University Hospital \\
  Freiburg, Germany \\
  \texttt{\href{mailto:angelo.d.ambrosio@uniklinik-freiburg.de}{\nolinkurl{angelo.d.ambrosio@uniklinik-freiburg.de}};
\href{mailto:a.dambrosioMD@gmail.com}{\nolinkurl{a.dambrosioMD@gmail.com}}} \\
   \And
    Hajo Grundmann
   \\
    Institute for Infection Prevention and Hospital Hygiene \\
    Freiburg University Hospital \\
  Freiburg, Germany \\
  \texttt{\href{mailto:hajo.grundmann@uniklinik-freiburg.de}{\nolinkurl{hajo.grundmann@uniklinik-freiburg.de}}} \\
   \And
    Tjibbe Donker
   \\
    Institute for Infection Prevention and Hospital Hygiene \\
    Freiburg University Hospital \\
  Freiburg, Germany \\
  \texttt{\href{mailto:tjibbe.donker@uniklinik-freiburg.de}{\nolinkurl{tjibbe.donker@uniklinik-freiburg.de}}} \\
  }

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{caption}
\usepackage{placeins}
\usepackage{amsmath}
\usepackage{lineno}
\linenumbers
\captionsetup[table]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\begin{document}
\maketitle

\begin{abstract}
The exponential growth of scientific production makes secondary
literature abridgements increasingly demanding. We introduce a new
open-source framework for systematic reviews that significantly reduces
time and workload for collecting and screening scientific literature.\\
The framework provides three main tools: 1) an automatic citation search
engine and manager that collects records from multiple online sources
with a unified query syntax, 2) a Bayesian, active machine learning,
citation screening tool based on iterative human-machine interaction to
increase predictive accuracy and, 3) a semi-automatic, data-driven query
generator to create new search queries from existing citation data
sets.\\
To evaluate the automatic screener's performance, we estimated the
median posterior sensitivity and efficiency {[}90\% Credible
Intervals{]} using Bayesian simulation to predict the distribution of
undetected potentially relevant records.\\
Tested on an example topic, the framework collected 17,755 unique
records through the citation manager; 766 records required human
evaluation while the rest were excluded by the automatic classifier; the
theoretical efficiency was 95.6\% {[}95.3\%, 95.7\%{]} with a
sensitivity of 100\% {[}93.5\%, 100\%{]}.\\
A new search query was generated from the labelled dataset, and 82,579
additional records were collected; only 567 records required human
review after automatic screening, and six additional positive matches
were found. The overall expected sensitivity decreased to 97.3\%
{[}73.8\%, 100\%{]} while the efficiency increased to 98.6\% {[}98.2\%,
98.7\%{]}.\\
The framework can significantly reduce the workload required to conduct
large literature reviews by simplifying citation collection and
screening while demonstrating exceptional sensitivity. Such a tool can
improve the standardization and repeatability of systematic reviews.
\end{abstract}

\keywords{
    Systematic review automation
   \and
    Citation management
   \and
    Online data collection
   \and
    Active machine learning
   \and
    Natural language processing
   \and
    Bayesian modeling
  }

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

the last decades (Larsen and Von Ins 2010; Bornmann and Mutz 2015). This
is especially true for biomedical research, a trend further accelerated
by the COVID-19 pandemic, thanks to faster article processing time by
publishers and the greater use of preprint databases (Aviv-Reuven and
Rosenfeld 2021; Horbach 2020; Hoy 2020). Consequently, it has become
harder for researchers and practitioners to stay up to date on the
latest findings in their field. Secondary research is of paramount
importance in this scenario in that it provides valuable summaries of
the latest research results; however, it is becoming ever more
challenging in terms of time and human resources required (Allen and
Olkin 1999; Borah et al. 2017; A. M. Cohen et al. 2010; Bastian,
Glasziou, and Chalmers 2010).\\

@@

particularly demanding (Babar and Zhang 2009). First, relevant published
research must be collected from scientific databases using appropriately
built search queries (retrieval phase); secondly, the scientific
citations collected must be screened, selecting only those that are
relevant to the topic (appraisal phase) (Bannach-Brown et al. 2019;
Tsafnat et al. 2014; Higgins et al. 2019).\\
Search queries construction is a complex task (Lefebvre et al. 2011;
Hammerstrøm et al. 2010), requiring both expertise in the scientific
field of interest and some knowledge of the database query languages.
The goal is to obtain a set of results that contains all relevant
articles (high sensitivity) while keeping the total number of records
low (high specificity), possibly focusing on the first at the expense of
the second (Hammerstrøm et al. 2010).\\

@@

record duplication (Marshall and Wallace 2019).\\

@@

small fraction are actually relevant (Lefebvre et al. 2011). It has been
estimated that labelling 10,000 publications can take up to 40 weeks of
work and that the average clinical systematic review takes 63 weeks to
complete (Bannach-Brown et al. 2019; Borah et al. 2017; Allen and Olkin
1999). A consequence of this is that systematic reviews are often
already out-of-date by the time they are published (E. M. Beller et al.
2013).\\

@@

has greatly maturated in the last years (Marshall and Wallace 2019; E.
Beller et al. 2018; Tsafnat et al. 2014). By applying natural language
processing (NLP), it is possible to transform free text into
quantitative features, with various levels of abstraction and
generalisation (Ananiadou and McNaught 2006; K. B. Cohen and Hunter
2008); using machine learning, such text-derived data can be used to map
and reproduce human judgment, automating the screening of citations
(Ikonomakis, Kotsiantis, and Tampakas 2005).\\

@@

the last years (Ananiadou et al. 2009; O'Mara-Eves et al. 2015; Tsafnat
et al. 2013; Jonnalagadda, Goyal, and Huffman 2015), and it is possible
foreseeable that it will become the standard approach in the field (E.
Beller et al. 2018), with many solutions already being implemented into
commercial or free-to-use tools (see Marshall and Wallace 2019, table
1).\\

@@

learning-based protocol (Settles 2009; Miwa et al. 2014), which utilises
a Bayesian framework to efficiently identify potentially relevant
documents that require human review while automatically screening-out
the vast majority of clearly non-relevant ones; the algorithm then
requires human review to increase classification accuracy iteratively.
Finally, we included a tool to generate new search queries based on an
already categorised citation data set, to identify relevant research
that manually-made queries may have possibly missed.\\

@@

microorganisms; the protocol is published in (Sadaghiani et al. 2020).\\

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{general-description}{%
\subsection{General description}\label{general-description}}

We built an R (R Core Team 2020) based framework to simplify two aspects
of systematic literature review: record acquisition and classification.
The code used to generate the results is available at
\url{https://github.com/AD-Papers-Material/BART_SystReviewClassifier},
while an updated and ready to use version of the framework is
distributed as an R package at
\url{https://github.com/bakaburg1/BaySREn}. The framework includes
several modules that communicate through intermediate outputs stored in
standard formats, making it possible for users to extend the framework
or easily integrate it with other tools in their pipeline. See
Supplemental Material S1 for an in-depth description of the framework
and how to use it.\\
The tasks carried out by the framework are grouped into ``sessions'',
which comprise obtaining scientific citation data (records) using a
search query and then labelling them as relevant (``positive'' in the
rest of the text) or not (``negative'') for the topic of interest with
the help of a machine learning engine (Fig. 1). The initial search query
should be built using domain knowledge, trying to achieve a high
relevant/non-relevant record ratio.\\
The framework can then generate a new data-driven query from this
labelled set to perform a new session to find records possibly missed by
the first query.\\

\begin{figure}
\includegraphics[width=1\linewidth]{methods_diagram} \caption{Figure 1. Framework's visual representation.}\label{fig:method_diagram}
\end{figure}

\hypertarget{record-acquisition-and-initial-labelling}{%
\subsection{Record acquisition and initial
labelling}\label{record-acquisition-and-initial-labelling}}

We built a set of tools to allow users to automatically search and
download citation data from three major scientific databases
(``sources''): Pubmed (\url{https://pubmed.ncbi.nlm.nih.gov/}), Web Of
Science (WOS, \url{https://apps.webofknowledge.com/}) and the Institute
of Electrical and Electronics Engineers (IEEE,
\url{https://ieeexplore.ieee.org/Xplore/home.jsp}). The framework
handles authorisation management for non-open databases like WOS and
IEEE. It is also possible to import previously downloaded records in the
framework; this is particularly useful for acquiring records from SCOPUS
(\url{https://www.scopus.com/}) and EMBASE databases
(\url{https://www.embase.com/}), for which a comprehensive API interface
was not easy to build. An extra manual search was also necessary for
Pubmed since the API and the web interface have different rule expansion
algorithms and return slightly different results ({`NCBI Insights :
Updated Pubmed e-Utilities Coming in April 2022!'}, n.d.). A short guide
on how to set up the framework for each database supported is available
in Supplemental Material S3.\\
The collected records are merged into a single database, resolving
duplicates and different formatting between sources. The records are
ordered according to the frequency of the positive query terms (e.g.,
not preceded by a \emph{NOT} modifier) in the title and abstract
(``simple query ordering'').\\
The researcher is then asked to label a subset of records to create the
``initial training set'' needed to start the automatic classification.
We recommend manually labelling the first 250 records (see
``hyperparameter optimisation'' later). Simple query ordering increases
the positivity rate in the initial training set (Wallace, Small, et al.
2010), leading to higher sensitivity during automatic classification
(Chawla, Japkowicz, and Kotcz 2004).

\hypertarget{text-feature-extraction}{%
\subsection{Text feature extraction}\label{text-feature-extraction}}

The framework models the relevance of a record based on the following
fields in the citation data: title, abstract, authors, keywords, MESH
terms (Lipscomb 2000). A range of Natural Language Processing (NLP)
techniques (Baeza-Yates, Ribeiro-Neto, et al. 1999; Marshall and Wallace
2019; Ananiadou and McNaught 2006) are employed to convert the textual
information in these fields into features for machine learning through a
bag-of-words approach (Marshall and Wallace 2019). Processing of free
text fields (title, abstract) includes: tokenisation (i.e., extracting
the terms), removal of common stopwords (i.e., sentence components
having no semantic value), part-of-speech filtering (only nouns,
adjectives, verbs and untagged terms are retained), and lemmatisation of
terms (i.e., reduction to their base grammatical form). Text processing
for authors, keywords and MESH terms identifies logical units (e.g.,
authors' full names, composite keywords) and extracts them.\\
Terms appearing in less than 5\% of the labelled documents are removed
from negative records. All terms in the positive set are kept to
increase sensitivity at the cost of specificity.\\
Some terms tend to co-appear in records (non-consecutive ngrams,
nc-ngrams), often carrying a particular meaning when they do co-occur.
To detect nc-ngrams, we generated a word network representation
(Francois Rousseau 2015) with edges occurring between terms with a
cosine similarity in terms of document co-occurrence \textgreater{} 0.5.
We extracted the maximal cliques in the network (Eppstein, Löffler, and
Strash 2010) representing highly correlated groups of terms; these
groups are added to the dataset as individual features. Only nc-ngrams
comprising a maximum of ten terms are kept.\\
A second network is built using a co-occurrence threshold of 0.9. In
this case, the cliques represent terms that always appear together and
can therefore be considered redundant (i.e., they do not need to be
considered separately). These terms are merged to increase computational
efficiency and reduce overfitting.\\
The output is a Document-Term Matrix (DTM), with \(N_d\) rows
representing the records (\(D_i\)), \(N_t\) terms column for the
\(t_{field}\) terms (divided by record field) and \({0,1}\) values
whether \(t_{field} \in D_i\). We also enriched the DTM with features
referencing the number of terms in each field to help the model scale
term importance based on the field length.

\hypertarget{label-prediction}{%
\subsection{Label prediction}\label{label-prediction}}

>>>>

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% https://www.biomedcentral.com/getpublished                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% https://miktex.org/                                             %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass[]{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath}
\RequirePackage[numbers]{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\usepackage[strings]{underscore} % CUSTOM: fixes problems with underscores in .bib

\usepackage{caption}
\usepackage{placeins}
\usepackage{amsmath}
\usepackage{lineno}
%\linenumbers
\captionsetup[table]{labelformat=empty}
\captionsetup[figure]{labelformat=empty}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \def\includegraphic{}
% \def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\def\url{} % CUSTOM: \url requires hyperref package, but loading it causes errors with the author formatting

\providecommand{\tightlist}{ % CUSTOM: pandoc uses \tightlist for lists, without loading the required package
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}
  }
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
  addressref={aff1},
    corref={aff1},
    email={angelo.d.ambrosio@uniklinik-freiburg.de,a.dambrosioMD@gmail.com}
]{\inits{A.D.}\fnm{Angelo} \snm{D'Ambrosio}}
\author[
  addressref={aff1},
    email={hajo.grundmann@uniklinik-freiburg.de}
]{\inits{H.G.}\fnm{Hajo} \snm{Grundmann}}
\author[
  addressref={aff1},
    email={tjibbe.donker@uniklinik-freiburg.de}
]{\inits{T.D.}\fnm{Tjibbe} \snm{Donker}}

% \author[
%   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
%   corref={aff1},                       % id of corresponding address, if any
% % noteref={n1},                        % id's of article notes, if any
%   email={jane.e.doe@cambridge.co.uk}   % email address
% ]{\inits{J.E.}\fnm{Jane E.} \snm{Doe}}
% \author[
%   addressref={aff1,aff2},
%   email={john.RS.Smith@cambridge.co.uk}
% ]{\inits{J.R.S.}\fnm{John R.S.} \snm{Smith}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{
    \orgdiv{Institute for Infection Prevention and Hospital Hygiene},
      \orgname{Freiburg University Hospital},
          \city{Freiburg},
      \cny{Germany}
  }

% \address[id=aff1]{%                           % unique id
%   \orgdiv{Department of Science},             % department, if any
%   \orgname{University of Cambridge},          % university, etc
%   \city{London},                              % city
%   \cny{UK}                                    % country
% }
% \address[id=aff2]{%
%   \orgdiv{Institute of Biology},
%   \orgname{National University of Sciences},
%   %\street{},
%   %\postcode{}
%   \city{Kiel},
%   \cny{Germany}
% }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{artnotes}
%%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
%\end{artnotes}

 \end{fmbox} % comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                           %%
%% The Abstract begins here                  %%
%%                                           %%
%% Please refer to the Instructions for      %%
%% authors on https://www.biomedcentral.com/ %%
%% and include the section headings          %%
%% accordingly for your article type.        %%
%%                                           %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{Background} %if any
The exponential growth of scientific production makes secondary
literature abridgements increasingly demanding. We introduce a new
open-source framework for systematic reviews that significantly reduces
time and workload for collecting and screening scientific literature.
\parttitle{Methods} %if any
The framework provides three main tools: 1) an automatic citation search
engine and manager that collects records from multiple online sources
with a unified query syntax, 2) a Bayesian, active machine learning,
citation screening tool based on iterative human-machine interaction to
increase predictive accuracy and, 3) a semi-automatic, data-driven query
generator to create new search queries from existing citation data
sets.\\
To evaluate the automatic screener's performance, we estimated the
median posterior sensitivity and efficiency {[}90\% Credible
Intervals{]} using Bayesian simulation to predict the distribution of
undetected potentially relevant records.\\
\parttitle{Results} %if any
Tested on an example topic, the framework collected 17,755 unique
records through the citation manager; 766 records required human
evaluation while the rest were excluded by the automatic classifier; the
theoretical efficiency was 95.6\% {[}95.3\%, 95.7\%{]} with a
sensitivity of 100\% {[}93.5\%, 100\%{]}.\\
A new search query was generated from the labelled dataset, and 82,579
additional records were collected; only 567 records required human
review after automatic screening, and six additional positive matches
were found. The overall expected sensitivity decreased to 97.3\%
{[}73.8\%, 100\%{]} while the efficiency increased to 98.6\% {[}98.2\%,
98.7\%{]}.\\
\parttitle{Conclusions} %if any
The framework can significantly reduce the workload required to conduct
large literature reviews by simplifying citation collection and
screening while demonstrating exceptional sensitivity. Such a tool can
improve the standardization and repeatability of systematic reviews.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
	\kwd{Systematic review automation}
	\kwd{Citation management}
	\kwd{Online data collection}
	\kwd{Active machine learning}
	\kwd{Natural language processing}
	\kwd{Bayesian modeling}
% \kwd{sample}
% \kwd{article}
% \kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
 %a  \end{fmbox} %uncomment this for two column layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                            %%
%% The Main Body begins here                  %%
%%                                            %%
%% Please refer to the instructions for       %%
%% authors on:                                %%
%% https://www.biomedcentral.com/getpublished %%
%% and include the section headings           %%
%% accordingly for your article type.         %%
%%                                            %%
%% See the Results and Discussion section     %%
%% for details on how to create sub-sections  %%
%%                                            %%
%% use \cite{...} to cite references          %%
%%  \cite{koon} and                           %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}      %%
%%                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%

\section*{Background}
\addcontentsline{toc}{section}{Background}

@@

the last decades \citep{larsen2010rate, bornmann2015growth}. This is
especially true for biomedical research, a trend further accelerated by
the COVID-19 pandemic, thanks to faster article processing time by
publishers and the greater use of preprint databases
\citep{aviv2021publication, horbach2020pandemic, hoy2020rise}.
Consequently, it has become harder for researchers and practitioners to
stay up to date on the latest findings in their field. Secondary
research is of paramount importance in this scenario in that it provides
valuable summaries of the latest research results; however, it is
becoming ever more challenging in terms of time and human resources
required
\citep{allen1999estimating, borah2017analysis, cohen2010evidence, bastian2010seventy}.\\

@@

particularly demanding \citep{babar2009systematic}. First, relevant
published research must be collected from scientific databases using
appropriately built search queries (retrieval phase); secondly, the
scientific citations collected must be screened, selecting only those
that are relevant to the topic (appraisal phase)
\citep{bannach2019machine, tsafnat2014systematic, higgins2019cochrane}.\\
Search queries construction is a complex task
\citep{lefebvre2011searching, hammerstrom2010searching}, requiring both
expertise in the scientific field of interest and some knowledge of the
database query languages. The goal is to obtain a set of results that
contains all relevant articles (high sensitivity) while keeping the
total number of records low (high specificity), possibly focusing on the
first at the expense of the second \citep{hammerstrom2010searching}.\\

@@

record duplication \citep{marshall2015systematic}.\\

@@

small fraction are actually relevant \citep{lefebvre2011searching}. It
has been estimated that labelling 10,000 publications can take up to 40
weeks of work and that the average clinical systematic review takes 63
weeks to complete
\citep{bannach2019machine, borah2017analysis, allen1999estimating}. A
consequence of this is that systematic reviews are often already
out-of-date by the time they are published
\citep{beller2013systematic}.\\

@@

has greatly maturated in the last years
\citep{marshall2015systematic, beller2018making, tsafnat2014systematic}.
By applying natural language processing (NLP), it is possible to
transform free text into quantitative features, with various levels of
abstraction and generalisation
\citep{ananiadou2006text, cohen2008getting}; using machine learning,
such text-derived data can be used to map and reproduce human judgment,
automating the screening of citations \citep{ikonomakis2005text}.\\

@@

the last years
\citep{ananiadou2009supporting, o2015using, tsafnat2013automation, jonnalagadda2015automating},
and it is possible foreseeable that it will become the standard approach
in the field \citep{beller2018making}, with many solutions already being
implemented into commercial or free-to-use tools \citep[see][table
1]{marshall2015systematic}.\\

@@

learning-based protocol \citep{settles2009active, miwa2014reducing},
which utilises a Bayesian framework to efficiently identify potentially
relevant documents that require human review while automatically
screening-out the vast majority of clearly non-relevant ones; the
algorithm then requires human review to increase classification accuracy
iteratively. Finally, we included a tool to generate new search queries
based on an already categorised citation data set, to identify relevant
research that manually-made queries may have possibly missed.\\

@@

microorganisms; the protocol is published in \citep{newis}.\\

@@

\section*{Methods}
\addcontentsline{toc}{section}{Methods}

\subsection*{General description}
\addcontentsline{toc}{subsection}{General description}

We built an R \citep{rstats2021} based framework to simplify two aspects
of systematic literature review: record acquisition and classification.
The code used to generate the results is available at
https://github.com/AD-Papers-Material/BART\_SystReviewClassifier, while
an updated and ready to use version of the framework is distributed as
an R package at https://github.com/bakaburg1/BaySREn. The framework
includes several modules that communicate through intermediate outputs
stored in standard formats, making it possible for users to extend the
framework or easily integrate it with other tools in their pipeline. See
Supplemental Material S1 for an in-depth description of the framework
and how to use it.\\
The tasks carried out by the framework are grouped into ``sessions'',
which comprise obtaining scientific citation data (records) using a
search query and then labelling them as relevant (``positive'' in the
rest of the text) or not (``negative'') for the topic of interest with
the help of a machine learning engine (Fig. 1). The initial search query
should be built using domain knowledge, trying to achieve a high
relevant/non-relevant record ratio.\\
The framework can then generate a new data-driven query from this
labelled set to perform a new session to find records possibly missed by
the first query.\\

\subsection*{Record acquisition and initial labelling}
\addcontentsline{toc}{subsection}{Record acquisition and initial
labelling}

We built a set of tools to allow users to automatically search and
download citation data from three major scientific databases
(``sources''): Pubmed (\url{https://pubmed.ncbi.nlm.nih.gov/}), Web Of
Science (WOS, \url{https://apps.webofknowledge.com/}) and the Institute
of Electrical and Electronics Engineers (IEEE,
\url{https://ieeexplore.ieee.org/Xplore/home.jsp}). The framework
handles authorisation management for non-open databases like WOS and
IEEE. It is also possible to import previously downloaded records in the
framework; this is particularly useful for acquiring records from SCOPUS
(\url{https://www.scopus.com/}) and EMBASE databases
(\url{https://www.embase.com/}), for which a comprehensive API interface
was not easy to build. An extra manual search was also necessary for
Pubmed since the API and the web interface have different rule expansion
algorithms and return slightly different results \citep{pubmedUpdate}. A
short guide on how to set up the framework for each database supported
is available in Supplemental Material S3.\\
The collected records are merged into a single database, resolving
duplicates and different formatting between sources. The records are
ordered according to the frequency of the positive query terms (e.g.,
not preceded by a \emph{NOT} modifier) in the title and abstract
(``simple query ordering'').\\
The researcher is then asked to label a subset of records to create the
``initial training set'' needed to start the automatic classification.
We recommend manually labelling the first 250 records (see
``hyperparameter optimisation'' later). Simple query ordering increases
the positivity rate in the initial training set
\citep{wallace2010active}, leading to higher sensitivity during
automatic classification \citep{chawla2004special}.

\subsection*{Text feature extraction}
\addcontentsline{toc}{subsection}{Text feature extraction}

The framework models the relevance of a record based on the following
fields in the citation data: title, abstract, authors, keywords, MeSH
terms \citep{lipscomb2000medical}. A range of Natural Language
Processing (NLP) techniques
\citep{baeza1999modern, marshall2015systematic, ananiadou2006text} are
employed to convert the textual information in these fields into
features for machine learning through a bag-of-words approach
\citep{marshall2015systematic}. Processing of free text fields (title,
abstract) includes: tokenisation (i.e., extracting the terms), removal
of common stopwords (i.e., sentence components having no semantic
value), part-of-speech filtering (only nouns, adjectives, verbs and
untagged terms are retained), and lemmatisation of terms (i.e.,
reduction to their base grammatical form). Text processing for authors,
keywords and MeSH terms identifies logical units (e.g., authors' full
names, composite keywords) and extracts them.\\
Terms appearing in less than 5\% of the labelled documents are removed
from negative records. All terms in the positive set are kept to
increase sensitivity at the cost of specificity.\\
Some terms tend to co-appear in records (non-consecutive ngrams,
nc-ngrams), often carrying a particular meaning when they do co-occur.
To detect nc-ngrams, we generated a word network representation
\citep{rousseau2015graph} with edges occurring between terms with a
cosine similarity in terms of document co-occurrence \textgreater{} 0.5.
We extracted the maximal cliques in the network
\citep{eppstein2010listing} representing highly correlated groups of
terms; these groups are added to the dataset as individual features.
Only nc-ngrams comprising a maximum of ten terms are kept.\\
A second network is built using a co-occurrence threshold of 0.9. In
this case, the cliques represent terms that always appear together and
can therefore be considered redundant (i.e., they do not need to be
considered separately). These terms are merged to increase computational
efficiency and reduce overfitting.\\
The output is a Document-Term Matrix (DTM), with \(N_d\) rows
representing the records (\(D_i\)), \(N_t\) terms column for the
\(t_{field}\) terms (divided by record field) and \({0,1}\) values
whether \(t_{field} \in D_i\). We also enriched the DTM with features
referencing the number of terms in each field to help the model scale
term importance based on the field length.

\subsection*{Label prediction}
\addcontentsline{toc}{subsection}{Label prediction}

We used a Bayesian Additive Regression Trees (BART) machine learning
``classification model'' \citep{chipman2010bart} \citep[in the
implementation of][]{kapelner2013bartmachine} to predict the probability
of a record being relevant, given the information coded into the
enriched DTM and the initial training set. We set up the BART model to
use 2,000 MCMC iterations (after 250 burn-in iterations) and 50 trees;
we used a \(k\) value of 2 to regularise extreme prediction and let the
model use missing fields in the DTM as features
\citep{kapelner2015prediction}. Positive records are oversampled ten
times to increase sensitivity \citep{batista2004study}.\\
The output is the expected value posterior predictive distribution for
each record (PPD in brief) describing the probability of it being
relevant (i.e., a positive match). An ensemble of ten models was fitted
to improve prediction stability by averaging the PPD between models
\citep{zhou2021ensemble, dietterich2000ensemble}.\\

To assign the labels, we employed an ``active learning'' approach
\citep{settles2009active, miwa2014reducing}, where a human reviews a
specific subset of predictions made by the machine, which is then
retrained on the manually reviewed dataset. This process is carried out
iteratively to reduce prediction uncertainty.\\
Label assignment is done through identification of an ``uncertainty
zone'', the construction of which is possible thanks to the Bayesian
nature of BART, which provides full PPDs instead of point-wise
predictions for each record.\\
To describe the process formally, we define:

\[\pi_i = \frac{1}{M}\sum_{j=1}^M Pr(L_i = \text{1}|DTM,m_j) \tag{Eq. 1}\]
